```{r}

library(tidyverse)
library(janitor)
library(lubridate)
library(here)
library(sf)
library(leaflet)
library(tsibble)
library(fable)
library(urca)
library(prophet)
library(htmltools)
# library(astsa)

```




```{r}
#for gg plots
beds <- read_csv(here("../../clean_data/beds_16_to_21.csv")) %>% 
  mutate(quarter = yearquarter(quarter))
beds

avg_beds <- read_csv(here("../../clean_data/avg_beds_16_to_21.csv")) %>% 
  mutate(quarter = yearquarter(quarter))
avg_beds

#for forecast
tsibble_beds <- read_csv(here("../../clean_data/tsibble_beds_16_to_21_percent.csv"))

tsibble_beds %>%
  mutate(quarter = yearquarter(quarter)) %>% 
  as_tsibble(index = quarter, key = healthboard) -> tsibble_beds
tsibble_beds


tsibble_wards <- read_csv(here("../../clean_data/tsibble_wards_16_to_21_percent.csv"))

tsibble_wards %>%
  mutate(quarter = yearquarter(quarter)) %>% 
  as_tsibble(index = quarter, key = healthboard) -> tsibble_wards
tsibble_wards
  

#for maps
beds_map <- st_read(here("../../clean_data/map_beds_4.geojson"))

beds_map

```

```{r}
beds_specialty <- sort(unique(beds$specialty_name))
beds_specialty
```



```{r}
beds
avg_beds

beds %>% 
  ggplot(aes(x = quarter, y = percentage_occupancy_by_board, group = healthboard, colour = healthboard)) +
  geom_line()+
  labs(title = "Percentage occupancy of beds for healthboards")

beds%>% 
  filter(healthboard == "Ayrshire and Arran") %>% 
  ggplot(aes(x = quarter, y = percentage_occupancy_by_board, group = healthboard)) +
  geom_line()+
  geom_line(show.legend = FALSE)+
  labs(title = "Percentage occupancy of beds for chosen healthboard",
       y = "Percent occupancy of beds",
       x = "Year/Quarter")+
  geom_hline(yintercept = 90, colour = "red")+
  theme_bw()+
  theme(panel.grid.minor.x = element_blank())+
  ylim(NA, 100)

beds%>% 
  filter(healthboard == "Ayrshire and Arran") %>% 
  filter(specialty_name == "All Acute") %>% 
  ggplot() +
  geom_point(aes(x = quarter, y = percentage_occupancy, group = quarter), alpha = 0.2)+
  geom_line(aes(x = quarter, y = percentage_occupancy_for_speciality, group = healthboard))+
  labs(title = "Percentage occupancy of beds for chosen healthboard",
       y = "Percent occupancy of beds",
       x = "Year/Quarter")+
  geom_hline(yintercept = 90, colour = "red")+
  theme_bw()+
  theme(panel.grid.minor.x = element_blank())+
  ylim(NA, 100)


beds %>% 
  distinct(specialty_name)
```
ggplot non interactive map

```{r}
beds_map %>% 
  ggplot()+
  geom_sf(aes(fill = mean_percent_occ_board))+
  labs()
```



```{r}

pal <- colorBin("YlOrRd", domain = beds_map$mean_percent_occ_board, bins = 7)

labels <- sprintf(
  "<strong>%s</strong><br/>%g &#x25 occupied beds on average <br> from Q4 2017 to Q3 2021",
  beds_map$healthboard, beds_map$mean_percent_occ_board
) %>% lapply(htmltools::HTML)



tag.map.title <- tags$style(HTML("
  .leaflet-control.map-title { 
    transform: translate(-50%,20%);
    position: fixed !important;
    left: 50%;
    text-align: center;
    padding-left: 10px; 
    padding-right: 10px; 
    background: rgba(255,255,255,0.75);
    font-weight: bold;
    font-size: 28px;
  }
"))

title <- tags$div(
  tag.map.title, HTML("Average percent occupancy of beds across NHS Sotland healthboards")
)  


map <- leaflet(beds_map) %>%
  addTiles() %>% 
  addPolygons(fillColor = ~pal(mean_percent_occ_board),
              color = "white",
              fillOpacity = 0.7,
              label = labels,
              weight = 2) %>% 
  addLegend(pal = pal,
            values = ~mean_percent_occ_board,
            opacity = 0.7,
            title = NULL,
            position = "bottomright") %>% 
  setView(lat = 58.0000, lng = -5.0000, zoom = 5)

map
```

Forecast data

```{r}

fit <- tsibble_beds %>% 
  model(
    snaive = SNAIVE(percentage_occupancy_by_board, level = 95)
  )

    # mean_model = MEAN(percentage_occupancy_by_board),
    # arima = ARIMA(percentage_occupancy_by_board)

forecast_1 <- fit %>%
  fabletools::forecast(h = 4)
forecast_1


forecast_1 %>% 
  filter(healthboard == "Borders") %>% 
  autoplot(tsibble_beds)+
  guides(colour = guide_legend(title = "Forecast"))
  

```


```{r}
avg_beds %>% 
  ggplot(aes(x = quarter, y = percent_wards_over_ninety, group = healthboard, colour = healthboard)) +
  geom_line()

beds %>% 
  filter(percent_wards_over_ninety != 0) %>% 
    filter(healthboard == "Fife") %>% 
  ggplot(aes(x = quarter, y = percent_wards_over_ninety, group = healthboard, colour = healthboard)) +
  geom_line(show.legend = FALSE)+
  labs(title = "Percent of wards in healthboard with over 90% bed occupancy",
       y = "Percent of wards with over 90% occupancy",
       x = "Year/Quarter")

```
```{r}
pal2 <- colorBin("YlOrRd", domain = beds_map$mean_wards_over_ninety, bins =5)

labels2 <- sprintf(
  "<strong>%s</strong><br/>%g &#x25 of wards over 90 <br> percent occupied from <br>Q4 2017 to Q3 2021",
  beds_map$healthboard, beds_map$mean_wards_over_ninety
) %>% lapply(htmltools::HTML)

map_2 <- leaflet(beds_map) %>%
  addTiles() %>% 
  addPolygons(fillColor = ~pal2(mean_wards_over_ninety),
              color = "white",
              fillOpacity = 0.7,
              label = labels2,
              weight = 2) %>% 
  addLegend(pal = pal2,
            values = ~mean_wards_over_ninety,
            opacity = 0.7,
            title = NULL,
            position = "bottomright")%>% 
  setView(lat = 58.0000, lng = -5.0000, zoom = 5)
map
```

```{r}

fit <- tsibble_wards %>% 
  model(
    snaive = SNAIVE(percent_wards_over_ninety),
    mean_model = MEAN(percent_wards_over_ninety),
    arima = ARIMA(percent_wards_over_ninety)
  )

forecast_2 <- fit %>%
  fabletools::forecast(h = 4)
forecast_2

forecast_2 %>% 
  filter(healthboard == "Ayrshire and Arran") %>% 
  autoplot(tsibble_wards)+
  guides(colour = guide_legend(title = "Forecast"))+
  labs(title = "What percentage occupancy of wards in chosen healthboard were at over 90% occupancy for the quarter?",
       x = "Year / Quarter",
       y = "Percent of wards with over 90% occupancy")

```

Forecasting difficult using 5 years of quarterly data, especially when there was a major disruption to normal operations (COVID) in the middle.




beds <- read_csv(here::here("../clean_data/beds_16_to_21.csv")) %>% 
  mutate(quarter = yearquarter(quarter))

beds_specialty <- sort(unique(beds$specialty_name))

beds_map <- st_read(here("../../clean_data/map_beds_4.geojson"))

pal <- colorBin("YlOrRd", domain = beds_map$mean_percent_occ_board, bins = 7)

labels <- sprintf(
  "<strong>%s</strong><br/>%g &#x25 occupied beds on average <br> from Q4 2017 to Q3 2021",
  beds_map$healthboard, beds_map$mean_percent_occ_board
) %>% lapply(htmltools::HTML)

pal2 <- colorBin("YlOrRd", domain = beds_map$mean_wards_over_ninety, bins =5)

labels2 <- sprintf(
  "<strong>%s</strong><br/>%g &#x25 of wards over 90 <br> percent occupied from <br>Q4 2017 to Q3 2021",
  beds_map$healthboard, beds_map$mean_wards_over_ninety
) %>% lapply(htmltools::HTML






  output$map_occ <- renderLeaflet({
  leaflet(beds_map) %>%
    addTiles() %>% 
    addPolygons(fillColor = ~pal(mean_percent_occ_board),
                color = "white",
                fillOpacity = 0.7,
                label = labels,
                weight = 2) %>% 
    addLegend(pal = pal,
              values = ~mean_percent_occ_board,
              opacity = 0.7,
              title = NULL,
              position = "bottomright") %>% 
    setView(lat = 58.0000, lng = -5.0000, zoom = 5)
})
  
  output$map_over_ninety <- renderLeaflet({
    leaflet(beds_map) %>%
    addTiles() %>% 
    addPolygons(fillColor = ~pal2(mean_wards_over_ninety),
                color = "white",
                fillOpacity = 0.7,
                label = labels2,
                weight = 2) %>% 
    addLegend(pal = pal2,
              values = ~mean_wards_over_ninety,
              opacity = 0.7,
              title = NULL,
              position = "bottomright")%>% 
    setView(lat = 58.0000, lng = -5.0000, zoom = 5)
  })
  
  output$beds_board_spec <- renderPlot({
    beds%>% 
    filter(healthboard == "Ayrshire and Arran") %>% 
    filter(specialty_name == "All Acute") %>% 
    ggplot() +
    geom_point(aes(x = quarter, y = percentage_occupancy, group = quarter), alpha = 0.2)+
    geom_line(aes(x = quarter, y = percentage_occupancy_for_speciality, group = healthboard))+
    labs(title = "Percentage occupancy of beds for chosen healthboard",
         y = "Percent occupancy of beds",
         x = "Year/Quarter")+
    geom_hline(yintercept = 90, colour = "red")+
    theme_bw()+
    theme(panel.grid.minor.x = element_blank())+
    ylim(NA, 100)
  })
  
  output$beds_over_ninety <- renderPlot({
    beds %>% 
    filter(percent_wards_over_ninety != 0) %>% 
    filter(healthboard == "Fife") %>% 
    ggplot(aes(x = quarter, y = percent_wards_over_ninety, group = healthboard, colour = healthboard)) +
    geom_line(show.legend = FALSE)+
    labs(title = "Percent of wards in healthboard with over 90% bed occupancy",
         y = "Percent of wards with over 90% occupancy",
         x = "Year/Quarter")
  })
  